---
title: "Reproducible Research: Peer Assessment 1"
output: github_document

---


## Loading and preprocessing the data

```{r, results='hide'}
library(Hmisc)
library(dplyr)
library(ggplot2)
if(!file.exists('activity.csv')){
        link <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
        download.file(link, "repdata%2Fdata%2Factivity.zip",method="curl")
        unzip("repdata%2Fdata%2Factivity.zip")
}

dat <- read.csv("activity.csv")
str(dat)
dim(dat)
dat$date <- as.Date(dat$date,"%Y-%m-%d")

```

## What is mean total number of steps taken per day?
```{r,fig.width=5,fig.height=5}
step.sum <- as.matrix(with(dat,by(steps,date,sum,na.rm=T)))
hist(step.sum,breaks = 30, xlab="Total number of steps taken per day",main="Total number of steps taken per day")
mean.step.sum <- round(mean(step.sum),2)
median.step.sum <- median(step.sum)
```
The mean and median of the total number of steps taken per day are `r mean.step.sum` and `r median.step.sum`  

## What is the average daily activity pattern?
```{r,fig.width=5,fig.height=5}
step.ave <- aggregate(dat$steps,list(dat$interval),mean,na.rm=T)
with(step.ave,plot(Group.1,x,type = "l",xlab="Interval",ylab="Average number of steps"))
max.strp <- step.ave[which.max(step.ave$x),]

```
On average across all the days in the dataset, `r max.strp[,1]` contains the maximum number of steps `r round(max.strp[,2],2)`  

## Imputing missing values

```{r,fig.width=5,fig.height=5}
na.num <- sum(is.na(dat$steps))
dat.fill <- dat
dat.fill$steps <- impute(dat.fill$steps,mean)
step.sum.data.fill <- as.matrix(with(dat.fill,by(steps,date,sum,na.rm=T)))
hist(step.sum.data.fill,breaks = 30, xlab="Total number of steps taken per day",main="Total number of steps taken per day")
mean.step.sum.data.fill <- round(mean(step.sum.data.fill),2)
median.step.sum.data.fill <- median(step.sum.data.fill)

```
-The total number of missing values in the dataset is `r na.num`.  
-The mean and median total number of steps taken per day are `r mean.step.sum.data.fill` and `r median.step.sum.data.fill` with the missing data filled in. 
-Compared with the first part of the assignment, the mean and median become lager.


## Are there differences in activity patterns between weekdays and weekends?
```{r,fig.width=5,fig.height=10}
weekday <- weekdays(dat.fill$date,abbreviate=T)
dat.fill$weekday <- "weekday"
dat.fill$weekday[weekday == "Sat" | weekday == "Sun"] <- "weekend"
table(dat.fill$weekday)
par(mfrow=c(2,1))
step.ave.data.fill <- aggregate(steps~weekday+interval,data=dat.fill, mean,na.rm=T)
with(step.ave.data.fill[step.ave.data.fill$weekday=="weekend",],plot(x=interval,y=steps,type = "l",xlab="interval",ylab="Number of Steps",main="Weekend"))
with(step.ave.data.fill[step.ave.data.fill$weekday=="weekday",],plot(x=interval,y=steps,type = "l",xlab="interval",ylab="Number of Steps",main="Weekdat"))
```